The pseudo-genetic view of science reframes evaluation from ``who is most cited'' to ``how much of a paper is inherited versus newly contributed.'' This distinction is useful for research assessment, literature mapping, and funding or review workflows that seek genuine conceptual advance.

\subsection{Interpretation of contribution}
A high contribution factor should not be read as universal quality, and a high inheritance factor should not be read as weakness. Consolidation papers, benchmarks, and systematic reviews can be scientifically valuable even when semantically close to prior work. Conversely, large residual novelty may reflect innovation, but may also indicate poor citation coverage, domain shift, or embedding noise. Contribution should therefore be interpreted alongside methodological rigor and empirical validity.

\subsection{Practical uses}
The framework supports several concrete applications:
\begin{itemize}
    \item \textbf{Lineage tracing}: identify dominant conceptual parents of influential papers.
    \item \textbf{Reviewer support}: flag papers whose claimed novelty appears inconsistent with inherited semantic mass.
    \item \textbf{Portfolio analysis}: quantify how programs or institutions balance incremental and exploratory research.
    \item \textbf{Field cartography}: reveal directional transfer of methods across disciplines.
\end{itemize}

\subsection{Methodological considerations}
Three caveats are central. First, embeddings compress complex texts, so results depend on representation quality and coverage of full methodological detail. Second, citation graphs are incomplete and noisy; missing references can inflate residual estimates. Third, decomposition under correlated parents is inherently non-identifiable without constraints, which is why uncertainty reporting is essential.

Despite these caveats, the framework offers a transparent and extensible basis for content-aware scientometrics. Its main strength is interpretability at both edge and node levels, enabling analyses that are difficult with purely relational or purely latent predictive models.
