\subsection{Overview}
The method has four stages: (1) construct a semantically embedded paper corpus, (2) build a time-respecting citation DAG, (3) estimate inheritance weights from cited parents to each focal paper, and (4) derive contribution and lineage statistics from the fitted decomposition.

\subsection{Construction of science space}
Let each paper $p_i$ be represented by title and abstract text. We encode this text using SciBERT to obtain an embedding $\mathbf{e}_i \in \mathbb{R}^d$. Embeddings are $\ell_2$-normalized to stabilize cosine geometry and reduce scale effects between documents of different length.

To improve robustness, we apply light preprocessing (Unicode normalization, citation-marker removal, and deduplication by identifier). We then compute corpus-level diagnostics (norm distribution, nearest-neighbor density, and year-stratified centroid drift) to verify embedding consistency before downstream estimation.

\subsection{Construction of citation DAG}
From metadata, each directed edge $(j \rightarrow i)$ denotes that paper $p_i$ cites paper $p_j$. We enforce temporal validity by removing edges where publication year$(p_j)$ exceeds publication year$(p_i)$. The resulting graph is acyclic after temporal filtering and transitive inconsistencies are logged for audit.

For each focal paper $p_i$, define its parent set $\mathcal{P}(i)$ as cited predecessors retained after filtering. Papers with fewer than $k_{\min}$ valid parents are kept in the corpus for descriptive analysis but excluded from inheritance regression to avoid unstable estimates.

\subsection{Inheritance inference}
For each eligible paper $p_i$, we model its embedding as
\begin{equation}
\mathbf{e}_i \approx \sum_{j \in \mathcal{P}(i)} w_{ij}\,\mathbf{e}_j + \mathbf{r}_i,
\end{equation}
where $w_{ij} \ge 0$ and $\mathbf{r}_i$ is a residual contribution vector.

Weights are estimated with non-negative constrained least squares and simplex regularization:
\begin{equation}
\min_{\mathbf{w}_i \ge 0}\; \left\|\mathbf{e}_i-\mathbf{E}_{\mathcal{P}(i)}\mathbf{w}_i\right\|_2^2 + \lambda\|\mathbf{w}_i\|_2^2
\quad\text{s.t.}\quad
\sum_{j \in \mathcal{P}(i)} w_{ij} \le 1,
\end{equation}
with $\mathbf{E}_{\mathcal{P}(i)}$ the parent-embedding matrix.

The non-negativity constraint enforces interpretable additive inheritance. The simplex bound prevents over-attribution when parent embeddings are highly collinear. The unassigned mass,
\begin{equation}
c_i = 1 - \sum_{j \in \mathcal{P}(i)} w_{ij},
\end{equation}
serves as a scalar \emph{contribution factor} (higher implies less explainability from cited parents). The vector residual
\begin{equation}
\mathbf{r}_i = \mathbf{e}_i-\sum_{j \in \mathcal{P}(i)} w_{ij}\mathbf{e}_j
\end{equation}
captures direction-specific novelty for downstream semantic inspection.

\subsection{Handling non-orthogonal parents}
Cited parents are typically correlated. To avoid unstable coefficient allocation, we combine three controls: (i) ridge regularization ($\lambda>0$), (ii) optional parent-pruning by semantic redundancy threshold, and (iii) bootstrap re-estimation over parent subsets to compute confidence intervals for $w_{ij}$. We report both point estimates and uncertainty to distinguish robust inheritance edges from ill-conditioned assignments.

\subsection{Derived metrics and analyses}
From fitted weights we compute:
\begin{itemize}
    \item \textbf{Edge inheritance strength} $w_{ij}$ for lineage tracing.
    \item \textbf{Paper contribution magnitude} $c_i$ and $\|\mathbf{r}_i\|_2$.
    \item \textbf{Lineage concentration} (entropy of $\mathbf{w}_i$), indicating whether inheritance is diffuse or dominated by a few parents.
    \item \textbf{Cross-field transfer} by aggregating inheritance mass across discipline labels.
\end{itemize}

These quantities support both ranking-style comparisons and structural analyses of knowledge evolution in the citation DAG.
