Research on scientific-publication analysis can be grouped into relational bibliometrics, semantic modeling, and hybrid science-of-science methods. Our framework builds on all three while targeting a specific gap: estimating inherited versus novel semantic content at the paper level.

\subsection{Relational bibliometrics and citation-network analysis}
Classical scientometrics treats citations as links in a graph and studies structure using count-based and network-based statistics. Citation totals, impact factor variants, and author-level indices provide simple, scalable proxies for influence, but they are coarse and field-sensitive. Network methods such as PageRank-style centrality, co-citation, bibliographic coupling, and community detection better capture structural position and diffusion pathways. Temporal extensions model growth, obsolescence, and burst dynamics across time slices.

These methods are valuable for mapping influence, yet they generally assume that each citation edge has comparable semantic meaning. In practice, citations can indicate foundational dependence, contrast, perfunctory acknowledgment, or methodological reuse. Without content, relational metrics cannot reliably distinguish these roles.

\subsection{Semantic representations of scientific text}
Semantic scientometrics introduces textual representations to recover meaning beyond graph topology. Early approaches relied on bag-of-words vectors and topic models; more recent work uses contextual embeddings from transformer models trained on scholarly corpora. Such representations support document similarity, clustering, trend analysis, entity linking, and literature retrieval.

A related line studies citation-context semantics by analyzing the sentence or paragraph surrounding citation mentions. This improves understanding of citation intent and rhetorical function. However, most semantic approaches still focus on similarity or classification and do not explicitly decompose a paper into inherited versus novel components.

\subsection{Hybrid text--graph approaches}
Hybrid methods combine citation structure with text features through graph neural models, heterogeneous networks, or joint embedding objectives. These systems improve tasks such as recommendation, field classification, and influence prediction by leveraging both topology and content. They demonstrate that semantic and relational signals are complementary.

Yet, even in hybrid settings, the output is often a latent score optimized for prediction tasks rather than an interpretable decomposition aligned with scientific contribution. This limits their use for transparent evaluation and policy-facing analysis.

\subsection{Paradigm shifts and novelty detection}
Work inspired by Kuhnian dynamics attempts to detect scientific revolutions through topic transitions, citation reconfiguration, and disruption indices. Novelty measures often rely on atypical combinations of references, semantic distance from prior work, or divergence from field norms. These methods offer useful indicators of change, but many are sensitive to parameterization and do not directly quantify what fraction of a paper is inherited from specific predecessors.

\subsection{Gap addressed by this study}
Across these literatures, no widely adopted framework simultaneously provides: (i) edge-level inheritance strength among cited parents, (ii) node-level residual novelty for each paper, and (iii) integration of both signals into a lineage-oriented citation DAG. The pseudo-genetic framework proposed here is designed to fill that gap with an interpretable, reproducible pipeline grounded in modern scientific-text embeddings.
