\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore sciento-metrics}

\begin{document}

\title{A Pseudo-Genetic Framework for Scientific Publications}

\author{Louis~George%
\thanks{Manuscript prepared April 26, 2026.}
}

\markboth{IEEE Transactions on Knowledge and Data Engineering,~Vol.~XX, No.~XX, Month~2026}%
{George: A Pseudo-Genetic Framework for Scientific Publications}

\maketitle

\begin{abstract}
This paper introduces a semantic scientometric framework that models scientific publications as evolving entities in a shared embedding space. Rather than treating all citations as equivalent, the framework decomposes each paper into two components: inherited content explained by cited predecessors and a residual contribution that captures novel authorial input. Using SciBERT embeddings and constrained regression over citation neighborhoods, we define an inheritance factor for each parent--child edge and a contribution factor for each paper. We then embed these estimates in a directed acyclic citation graph to recover knowledge lineages and characterize local and global evolution patterns. On a corpus of approximately 5,000 open-access papers, the method produces interpretable inheritance profiles, improves discrimination between derivative and novel papers relative to count-based metrics, and reveals cross-disciplinary transfer patterns that are obscured by conventional bibliometrics. The framework provides a practical foundation for content-aware evaluation of scientific contribution and for studying how ideas accumulate, recombine, and diverge over time.
\end{abstract}

\begin{IEEEkeywords}
Scientometrics, citation analysis, semantic embeddings, scientific contribution, knowledge graphs, novelty detection.
\end{IEEEkeywords}

\section{Introduction}
\input{sections/intro.tex}

\section{Related Work}\label{sec:related}
\input{sections/related_work.tex}

\section{Method}\label{sec:method}
\input{sections/method.tex}

\section{Experiments}\label{sec:experiments}
\input{sections/experiments.tex}

\section{Discussion}\label{sec:discussion}
\input{sections/discussion.tex}

\section{Limitations \& Future Work}\label{sec:lim_future}
\input{sections/future.tex}

\section{Conclusion}\label{sec:conclusion}
\input{sections/conclusion.tex}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
